---
title: "Sample Breaking News Story"
author: "Riya Sharma"
format:
  html:
    self-contained: true
    code-fold: true
    code-tools: true
---

```{r setup, include=FALSE, warning=FALSE, message=FALSE}

library(tidyverse)
library(tigris)
library(ggplot2)
library(htmltools)
library(janitor)
library(here)

```

```{r, include=FALSE, warning=FALSE, message=FALSE}

preg = read_csv("/cloud/project/NationalAndStatePregnancy_PublicUse.csv")

head(preg)

```

```{r, include=FALSE}
preg2017 = preg %>% 
  filter(year==2017) %>%
  summarize(mean_abortion =
            janitor::round_half_up(mean(abortionratetotal), 2),
            mean_2024 = janitor::round_half_up(mean(abortionrate2024), 2))
```

# Breaking News: Access to Abortion Decreases Across the Country

\* *This is an example breaking news story regarding abortion access that uses inline R code \**

On Thursday, May 4, North Carolina approved [legislation](https://www.nytimes.com/2023/05/04/us/abortion-ban-north-carolina.html?searchResultPosition=4) banning almost all abortions after 12 weeks. North Carolina is following the lead of Florida, North Dakota, and 12 other states who enacted abortion bans after the Supreme Court [overturned](https://www.nytimes.com/interactive/2022/06/24/us/politics/supreme-court-dobbs-jackson-analysis-roe-wade.html) Roe v. Wade in June.

The North Carolina Senate voted 29-20 to approve the ban. However, the governor has 10 days to decide whether to veto it or not.

The Republican party can still override the veto if they receive an adequate number of votes.

DeSantis's strict legislation bans abortions after 6 weeks in [Florida](https://www.nytimes.com/2023/04/13/us/florida-six-week-abortion-ban.html?searchResultPosition=32), even though one in three people do not know they are pregnant until after this period. Hailed as a refuge for those in the Deep South seeking abortions, Florida is now scaling back on its protections.

This news comes as an increasing number of restrictions have been placed on abortion access. On April 8, a Texas federal judge ruled that [mifepristone](https://www.fda.gov/drugs/postmarket-drug-safety-information-patients-and-providers/questions-and-answers-mifepristone-medical-termination-pregnancy-through-ten-weeks-gestation), a key medication used to end pregnancy within 10 weeks, violated a federal rule.

Though the Supreme Court temporarily [preserved](https://www.cbsnews.com/news/supreme-court-abortion-pill-access-mifepristone/) access to these pills, the future of reproductive health hangs in the balance, especially as states like Alabama, Texas, and Missouri uphold pre-Roe trigger [bans](https://reproductiverights.org/maps/abortion-laws-by-state/) that include civil and criminal penalties for seeking and performing abortions.

An average of `r preg2017$mean_abortion` percent of people receive abortions nationwide, with the highest average, `r preg2017$mean_2024` percent, being among women aged 20 to 24.

With reproductive health under threat, Americans are mobilizing to maintain protections. In response to Florida's ban, a [coalition](https://www.nytimes.com/2023/05/05/us/politics/florida-abortion-ballot-ban-six-weeks.html?searchResultPosition=7) of organizers is aiming to reduce "government interference with abortion" by placing a measure for abortion access on the ballot next year.

------------------------------------------------------------------------

*You can see the code I used for the inline statistics by viewing the breaking.qmd page in my github repository [here](https://github.com/raashar/rsharma-site).*
