---
title: "Election Project (Mapmaking)"
author: "Riya Sharma, based on a lecture by Aaron Kessler"
format:
  html:
    self-contained: true
    code-fold: true
    code-tools: true
---


```{r setup, include=FALSE, warning=FALSE, message=FALSE}
library(sf)
library(mapview)
library(tmap)
library(tmaptools)
library(htmltools)
library(rmapshaper)
library(tigris)
library(tidycensus)
library(writexl)
library(jsonlite)
options(tigris_class = "sf")
library(tidyverse)
library(janitor)
library(kableExtra)
library(here)
options(scipen = 999)
options(stringsAsFactors = FALSE)


#load saved joined data file from previous script
joined_vacomparison <- readRDS(here("processed_data", "joined_vacomparison.rds"))



```


# Comparing Election Results for Virginia Governor and President



```{r}

head(joined_vacomparison)


```
```{r}
joined_vacomparison = joined_vacomparison %>%
  mutate(
    mc_overperform = pct_mcauliffe - biden_pct,
    dem_gov_margin = pct_mcauliffe - pct_youngkin,
    rep_gov_margin = pct_youngkin - pct_mcauliffe,
    dem_pres_margin = biden_pct - trump_pct,
    rep_pres_margin = trump_pct - biden_pct
  )
```

## Visualizing the areas in which Democrats won by a higher margin (Governor)
 
We'll use tmap and get the geometries for VA

```{r}
va_counties_geo <- tigris::counties(state = "VA", resolution = "20m", cb = TRUE)

va_counties_geo$NAMELSAD = toupper(va_counties_geo$NAMELSAD)

va_elections_sf = va_counties_geo %>%
  left_join(. , joined_vacomparison, by=c("NAMELSAD"="locality"))

va_elections_sf = va_elections_sf %>% mutate(
    dem_gov_margin = janitor::round_half_up(dem_gov_margin, 2)
  )

va_elections_sf = va_elections_sf %>% mutate(
    dem_pres_margin = janitor::round_half_up(dem_pres_margin, 2)
  )
```

Now, we'll actually make a map of the Democratic margins for Governor

```{r}

mylabel1 = glue::glue("<strong>{va_elections_sf$NAMELSAD}</strong><br />
                      Democratic Margin for Governor: {va_elections_sf$dem_gov_margin}<br />") %>% 
  lapply(htmltools::HTML)

map1 = mapview(va_elections_sf, zcol = "dem_gov_margin",
               col.regions = RColorBrewer::brewer.pal(9, "PuBuGn"),
               alpha.regions = 1,
               label = mylabel1,
               legend = F)
map1
```

This map helps us visualize how different counties vote very differently. While some pockets in the state's center, North, and East regions overwhelmingly voted for the Democratic candidate for Governor, a majority chose the Republican candidate.

## Visualizing the areas in which Democrats won by a higher margin (President)

Now, we'll do the same and make a map of the Democratic margins for President.

```{r}

mylabel2 = glue::glue("<strong>{va_elections_sf$NAMELSAD}</strong><br />
                      Democratic Margin for President: {va_elections_sf$dem_pres_margin}<br />") %>% 
  lapply(htmltools::HTML)

map2 = mapview(va_elections_sf, zcol = "dem_pres_margin",
               col.regions = RColorBrewer::brewer.pal(9, "PuBuGn"),
               alpha.regions = 1,
               label = mylabel2,
               legend = F)
map2
```

Here, we see the results largely mirror those for Governor. However, we see some counties (in the Southeast, particularly) have a greater Democratic margin for President than for Governor. 

