---
title: "Analysis Walkthrough Example"
author: "Riya Sharma, based on a lecture by Aaron Kessler"
format:
  html:
    self-contained: true
    code-fold: true
    code-tools: true

---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(tigris)
library(sf)
library(tidycensus)
library(htmltools)
library(janitor)
library(here)
library(mapview)
library(leafsync)
library(leaflet.extras2)
options(tigris_class = "sf")


```

# Interactive Maps with MapView Package

We'll be using the tidycensus package to pull both census data, as well as geospatial boundaries. 
Let's quickly review how we use it.

```{r echo=FALSE}

# uncomment to run, then recomment it out so you don't run it every time

census_api_key("2a6f8c21a30d3024e038d67d7d4eba647dc79cd4", install=TRUE, overwrite = TRUE)


```

Here, we choose variables we want
```{r}
#chose variables we want
myvars <- c(totalpop = "B01003_001",
            medincome = "B19013_001",
            medage = "B01002_001"
)
```

Now, we pull for GA counties
```{r}
#pull for GA counties
ga_counties_withgeo <- get_acs(geography = "county",
                       variables = c(myvars),
                       state = "GA",
                       output = "wide",
                       geometry = TRUE)

ga_counties_withgeo
```

We can also get all counties in the US
```{r}
#all counties in the US
all_counties_withgeo <- get_acs(geography = "county",
                       variables = c(myvars),
                       output = "wide",
                       geometry = TRUE)

all_counties_withgeo
```

Then, we remove MOE columns - they all end with "M"
```{r}
#remove MOE columns - they all end with "M"
ga_counties_withgeo <- ga_counties_withgeo %>%
  select(-ends_with("M"))

ga_counties_withgeo
```

...and also remove that trailing "E"
```{r}
#remove that trailing "E"
colnames(ga_counties_withgeo) <- sub("E$", "", colnames(ga_counties_withgeo)) # $ means end of string only

ga_counties_withgeo
```

Mapping GA counties with mapview

```{r}

mapview(ga_counties_withgeo, zcol = "medincome")

mapview(ga_counties_withgeo, zcol = "medage")

```


Customize colors
```{r}

mapview(ga_counties_withgeo, zcol = "medincome", 
         col.regions = RColorBrewer::brewer.pal(9, "Greens"), 
         alpha.regions = 1)



```
This map’s dark background appeared automatically, because mapview determined the map included a lot of light colors. You can turn off that feature.

```{r}

mapviewOptions("basemaps.color.shuffle" = FALSE)


```


```{r}

mapview(ga_counties_withgeo, zcol = "medincome", 
         col.regions = RColorBrewer::brewer.pal(9, "Greens"), 
         alpha.regions = 1)



```

Two maps together
```{r}

map_income <- mapview(ga_counties_withgeo, zcol = "medincome", 
         col.regions = RColorBrewer::brewer.pal(9, "Greens"), 
         alpha.regions = 1)

map_age <- mapview(ga_counties_withgeo, zcol = "medage", 
         col.regions = RColorBrewer::brewer.pal(9, "Greens"), 
         alpha.regions = 1)


```


```{r}
# two maps together
sync(map_income, map_age)


```

Side-by-side slider to compare the two, from the leaflet.extras2 package 
```{r}

map_income | map_age


```
To turn off legends, hover text, popups


```{r}

mapview(ga_counties_withgeo, zcol = "medincome", 
         col.regions = RColorBrewer::brewer.pal(9, "Greens"), 
         alpha.regions = 1,
         legend = FALSE, 
         label = FALSE, 
         popup = FALSE)


```

Customize labels with the glue package
```{r}

mylabel <- glue::glue("{ga_counties_withgeo$NAM} {ga_counties_withgeo$medincome}")


mapview(ga_counties_withgeo, zcol = "medincome", 
         col.regions = RColorBrewer::brewer.pal(9, "Greens"), 
         alpha.regions = 1,
         label = mylabel)

# makes label of county name and median income

```

Customize popups
```{r}
mypopup <- glue::glue("<strong>{ga_counties_withgeo$NAM}</strong><br />
                      Total Population: {ga_counties_withgeo$totalpop}<br />
                      Median Income: {ga_counties_withgeo$medincome}") %>% 
  lapply(htmltools::HTML)

# mylabel <- glue::glue("{all_data$State} {all_data$PctChange10_20}%") %>% lapply(htmltools::HTML)

```


```{r}

head(mypopup)

```


```{r}

mapview(ga_counties_withgeo, zcol = "medincome", 
         col.regions = RColorBrewer::brewer.pal(9, "Greens"), 
         alpha.regions = 1,
         popup = mypopup,
        label = mylabel)

```


